---
title: "DBI an Introduction"
author: "Wayne McClure"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

##so it begins

```

## Background 

One of the challenges in any programming language is redundancy. Particularly code that solves a common problem. If a mahalanobis distance algorithm is needed, the process for programmatically approaching the statistical method behind a mahalanobis distance and any subsequent derivations, such as graphs, can all be readily done with oine skilled enough in the statistical and programming knowledge to write from scratch
However, this type of analysis is common enough. Enough so that the community in R has created a package, a collection of code/functions stored in a central repository.[More can be seen here](https://www.tutorialspoint.com/r/r_packages.htm)

A great quote from Jenny Bryan
>Of course someone has to write loops. It doesn’t have to be you.

This is the general premise of Packages. However, like any quest gathering items and *packages* not everything you might need is contained in the same place and library.

```{r libraries}
library(dm)
library(DBI)
library(rquery)
library(odbc)
```

## The proposal and packages

The above packages all bring something to the table for proper data analysis of a potentially unknown database. 

My final project will bring together the unique functions from each into a single cohesive package allowing for the intelligent query, exploration, and connection to known databases.

Our first useful tool is the DBI and DM packages to create a ODBC connection to an existing MySQL Server database:

```{r, echo=TRUE}

con <- dbConnect(RMySQL::MySQL(), 
                             dbname = "employees", 
                             host = "192.168.50.71", 
                             port = 49153,
                             user = "root",
                             password = "Camplovers01")


```

The components of most us to us in the DM package are the `dm_from_src()` fucntion which passes the connection object, we just created above, into an argument that then can be used to pull in data from tables
```{r, echo=TRUE}
#> ── Table source ───────────────────────────────────────────────────────────
#> src:  mysql  
#> ── Metadata ───────────────────────────────────────────────────────────────
#> Tables: `departments`, `dept_emp`, `dept_manager`, `employees`, `salaries`, … (6 total)
#> Columns: 57

```


#The other components
Other areas where this final packakage will suceed will include visualization of the schemas using a ERD diagrams using functions from the RQuery library to bring in simple select statements to bring in tables as data objects.